Cerebral oxygenation manuscript figures

### 2022-02-14 final_figures.Rmd
export to pdf, run additional brain_swell models

---
title: "Cerebral oxygenation manuscript figures"
date: "`r Sys.Date()`"
output: html_document
---

@@

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10)

@@

### Fig 1. Hb_oxy and Hb_tot in the brain and muscle

```{r fig1}

@@

### Fig 2. Hb_oxy alpha and Hb_tot alpha in the brain and muscle

```{r fig2}

@@

### Fig 3. Increased cerebral_hb_tot and decreased cerebral_hb_tot_alpha are distinct clinical features of cerebral malaria

```{r fig3_data}

@@

#### A. Correlations of variables within subject groups

```{r fig3a, fig.show = "hold", out.width = "50%", fig.height = 12, fig.width = 15}

@@

#### B. Principal component analysis

```{r fig3b}

@@

#### C. Coefficients

```{r model, eval = FALSE}

@@

table3c <- hbox_final %>%
  tidy(exponentiate = TRUE) %>%
  filter(p.value < 0.05) %>% 
  left_join(df_CI, by = "term") %>%
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2)),
         conf.low = ifelse(`2.5 %` < 10^-2, scientific(`2.5 %`, digits = 3), round(`2.5 %`, 2)),
         conf.high = ifelse(`97.5 %` < 10^-2, scientific(`97.5 %`, digits = 3), round(`97.5 %`, 2))) %>%
  mutate(p.value = scientific(p.value, digits = 3),
         statistic = round(statistic, 2)) %>%
  unite("95% CI", c(conf.low, conf.high), sep = ", ") %>%
  dplyr::select(c(term, estimate, `95% CI`, statistic, p.value)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate",
                "t-statistic" = "statistic", "p-value" = "p.value") %>%
  as.data.frame()

@@

#### D. ROC curve of final model

```{r fig3d_plot, fig.height = 8}

@@

@@

### Fig 4. An increase in cerebral_hb_tot is significantly associated with an increase in brain swelling score

@@

  geom_smooth(method = "lm", lty = 2, color = "#696969", alpha = 0.2) +
  annotate("text", label = model, x = 25, y = 8, hjust = 0, vjust = 1, size = 5) +
  annotate("text", label = r2, x = 25, y = 7.5, hjust = 0, vjust = 1, size = 5) +
  annotate("text", label = "p-values", x = 25, y = 7, hjust = 0, vjust = 1, size = 5) +
  annotate("text", label = model_p_val, x = 35, y = 6.75, hjust = 0, vjust = 1, size = 4) +
  annotate("text", label = hb_tot_p_val, x = 35, y = 6.5, hjust = 0, vjust = 1, size = 4) +

@@

@@

@@

#### F. Simulation results

```{r fig_s1f}

@@

### SFig 2. Logistic regression curves

```{r fig_s2, fig.height = 12}

@@

**Figure S2.** Plotted are the individual logistic regression curves of individual variables tested for their discriminatory power between UM and CM. On the y-axis, 1 represents CM while 0 represents UM. A steeper curve indicates higher discriminatory power.

>>>>

---
title: "Cerebral oxygenation manuscript figures"
date: "`r Sys.Date()`"
output: pdf_document
---

@@

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "", fig.height = 6, fig.width = 10, 
                      fig.align = 'center')

@@

### Fig 1. Hb_oxy and Hb_tot in the brain and muscle
```{r fig1}

@@

```
 \newpage

### Fig 2. Hb_oxy alpha and Hb_tot alpha in the brain and muscle
```{r fig2}

@@

```
 \newpage

### Fig 3. Increased cerebral_hb_tot and decreased cerebral_hb_tot_alpha are distinct clinical features of cerebral malaria
```{r fig3_data}

@@

#### A. Correlations of variables within subject groups
```{r fig3a, fig.show = "hold", out.width = "50%", fig.height = 12, fig.width = 15}

@@

```
\newpage

#### B. Principal component analysis
```{r fig3b}

@@

```
\newpage

#### C. Coefficients
```{r model, eval = FALSE}

@@

table3c <- hbox_final %>%
  tidy(exponentiate = TRUE) %>%
  filter(p.value < 0.05) %>% 
  left_join(df_CI, by = "term") %>%
  mutate(estimate = ifelse(estimate < 10^-2, scientific(estimate, digits = 3), round(estimate, 2)),
         conf.low = ifelse(`2.5 %` < 10^-2, scientific(`2.5 %`, digits = 3), round(`2.5 %`, 2)),
         conf.high = ifelse(`97.5 %` < 10^-2, scientific(`97.5 %`, digits = 3), round(`97.5 %`, 2))) %>%
  mutate(p.value = scientific(p.value, digits = 3),
         statistic = round(statistic, 2)) %>%
  unite("95% CI", c(conf.low, conf.high), sep = ", ") %>%
  dplyr::select(c(term, estimate, `95% CI`, statistic, p.value)) %>% 
  dplyr::rename("Term" = "term", "Odds ratio" = "estimate",
                "t-statistic" = "statistic", "p-value" = "p.value") %>%
  filter(Term != "(Intercept)") %>% 
  as.data.frame()

@@

```
**note: for the OR, denominator is probability of patient having CM; numerator is p patient having UM

\newpage

#### D. ROC curve of final model
```{r fig3d_plot, fig.height = 8}

@@

```
\newpage

### Fig 4. An increase in cerebral_hb_tot is significantly associated with an increase in brain swelling score
```{r brainswell_model_comparison, eval = FALSE}

# remove cerebral_hb_tot outliers
  df_hbox_hbtot_outliers <- df_hbox %>% 
    mutate(Q1_hbtot = quantile(cerebral_hb_tot, 0.25, na.rm = TRUE),
           Q3_hbtot = quantile(cerebral_hb_tot, 0.75, na.rm = TRUE),
           IQR_hbtot = Q3_hbtot - Q1_hbtot,
           outlier_hbtot = case_when(

lm1 <- lm(brain_swell ~ cerebral_hb_tot, data = df_hbox_hbtot_outliers)
lm2 <- lm(brain_swell ~ cerebral_hb_tot + hct, data = df_hbox_hbtot_outliers)
lm3 <- lm(brain_swell ~ cerebral_hb_tot + cerebral_hb_tot_alpha2, data = df_hbox_hbtot_outliers)
lm4 <- lm(brain_swell ~ cerebral_hb_tot + cerebral_hb_tot_alpha2 + hct, data = df_hbox_hbtot_outliers)
lm5 <- lm(brain_swell ~ cerebral_hb_oxy + cerebral_hb_oxy_alpha2 + cerebral_hb_tot, data = df_hbox_hbtot_outliers)

# evaluate parameters for all models
brain_swell_models <- list(lm1, lm2, lm3, lm4, lm5)

df_brain_swell_models <- tibble(formula = as.character(),
                                term = as.character(),
                                aic = as.numeric(),
                                r2 = as.numeric(),
                                adj_r2 = as.numeric(),
                                model_p_val = as.numeric(),
                                estimate = as.numeric(),
                                std_error = as.numeric(),
                                statistic = as.numeric(),
                                term_p_val = as.numeric())

for (i in 1:length(brain_swell_models)) {

  formula <- brain_swell_models[[i]] %>% .$call
  formula <- gsub(".*=", "", formula)[[2]]

  aic <- brain_swell_models[[i]] %>% AIC
  r2 <- brain_swell_models[[i]] %>% summary %>% .$r.squared
  adj_r2 <- brain_swell_models[[i]] %>% summary %>% .$adj.r.squared
  f <- brain_swell_models[[i]] %>% summary %>% .$fstatistic
  p <- pf(f[1], f[2], f[3], lower.tail = FALSE)
  coef <- brain_swell_models[[i]] %>% tidy %>% filter(term != "(Intercept)") %>% dplyr::rename("term_p_val" = "p.value")

  df_tmp <- tibble(formula = formula,
                   term = coef$term,
                   aic = aic,
                   r2 = r2,
                   adj_r2 = adj_r2,
                   model_p_val = p) %>%
    left_join(coef) %>%
    clean_names

  df_brain_swell_models <- df_brain_swell_models %>% 
    bind_rows(df_tmp) %>% 
    relocate(term, .before = estimate)

df_brain_swell_models %>% write.csv(file = "outputs/brain_swell_models.csv")

@@

@@

  geom_smooth(method = "lm", lty = 2, color = "#696969", alpha = 0.2) +
  # annotate("text", label = model, x = 25, y = 8, hjust = 0, vjust = 1, size = 5) +
  # annotate("text", label = r2, x = 25, y = 7.5, hjust = 0, vjust = 1, size = 5) +
  # annotate("text", label = "p-values", x = 25, y = 7, hjust = 0, vjust = 1, size = 5) +
  # annotate("text", label = model_p_val, x = 35, y = 6.75, hjust = 0, vjust = 1, size = 4) +
  # annotate("text", label = hb_tot_p_val, x = 35, y = 6.5, hjust = 0, vjust = 1, size = 4) +

@@

```
**Figure 4.** model: brain_swell ~ cerebral_hb_tot + hematocrit; adjusted R^2: 0.25; model p-value: 0.007; cerebral_hb_tot p-value: 0.003

\newpage

@@

```
 \newpage

@@

```
 \newpage

#### F. Simulation results
```{r fig_s1f}

@@

```
 \newpage

### SFig 2. Logistic regression curves
```{r fig_s2, fig.height = 12}

@@

```
**Figure S2.** Plotted are the individual logistic regression curves of individual variables tested for their discriminatory power between UM and CM. On the y-axis, 1 represents CM while 0 represents UM. A steeper curve indicates higher discriminatory power.
